\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-05-17 podc16/paper2.tex
Working on the PODC camera ready.

@@

Most agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader process which enforces an ordering on commands during fault-free periods.
In practice, the unique leader can easily become a performance bottleneck for the system.
Therefore, recently proposed agreement algorithms like Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
Mencius rotates the leader among the processes but its structure remains close to MultiPaxos. 
However, EPaxos and Alvin are based on a novel algorithmic idea first introduced in EPaxos.

@@

On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions in varied environments such as clusters or wide area networks, and on varied metrics such as agreement latency, impact of failures or conflicts, tolerance to slow processes, load
balancing, size of quorums, etc.
We have formalized\footnote{\url{http://losa.fr/leaderless}} and model-checked our generic leaderless algorithm with the TLA$^+$ toolbox.

@@

We consider a set of processes $P$ which are subject to crash-stop faults and which communicate by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{GC-Propose}(c)$, must produce responses of the form $\opn{GC-Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
We say that two commands \emph{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ if and only if there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$. Moreover, $\bot\sqsubseteq \sigma$ for any
$\sigma$. Two c-structs are compatible when they have a
common upper bound \todo{that is constructible from \ldots}. A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$. The consistency property of generalized consensus requires that any two decided c-structs be compatible.
We refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus, and we use the same notation.

>>>>

\CopyrightYear{2016}
\setcopyright{rightsretained}
\conferenceinfo{PODC'16}{July 25-28, 2016, Chicago, IL, USA}
\isbn{978-1-4503-3964-3/16/07}
\doi{http://dx.doi.org/10.1145/2933057.2933072}

@@

Most agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader process which enforces an ordering on commands during fault-free periods.
In practice, the unique leader can easily become a performance bottleneck for the system.
Therefore, recently proposed agreement algorithms like S-Paxos~\cite{BielyETAL12SpaxosOffloadingLeaderHighThroughputState}, Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
S-Paxos offloads the computation that the leader has to perform to other replicas, while Mencius rotates the leader among the processes; the structure of both algorithms remains close to MultiPaxos. 
However, EPaxos and Alvin are based on a novel algorithmic idea first introduced in EPaxos.

@@

On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions in varied environments such as clusters or wide area networks, and on varied metrics such as agreement latency, impact of failures or conflicts, tolerance to slow processes, load
balancing, size of quorums, etc.
We have formalized\footnote{\url{http://losa.fr/leaderless}} and model-checked our generic leaderless algorithm using the TLA$^+$ toolbox.

@@

We consider a set of processes $P$ which are subject to crash-stop faults and which communicate by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{GC-Propose}(c)$, must produce responses of the form $\opn{GC-Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
We say that two commands \emph{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ if and only if there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$. Moreover, $\bot\sqsubseteq \sigma$ for any
$\sigma$. 
A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$. 
Two c-structs are compatible when they have a common upper bound that is constructible from the commands contained in the two c-structs. 
The non-triviality property of generalized consensus requires that any decided c-struct be of the form $\bot\bullet \mathcal{D}$, where $\mathcal{D}$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
if a command keeps beeing proposed, then a c-struct $\sigma$ containing the command is eventually decided.
We refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus, and we use the same notation.
